Containers and R: the Rockerverse and beyond

### 2019-10-08 manuscript.tex
draft addition of docker and rocker intro

\emph{Do we need a generic intro here?} DE: Yes, and happy to take a
stab.

@@

\address{%
Dirk Eddelbuettel\\
University of Illinois at Urbana-Champaign\\
\\
}
\href{mailto:dirkd@eddelbuettel.com}{\nolinkurl{dirkd@eddelbuettel.com}}

>>>>

Docker, an application and service provide by the eponymous company, has
in just a few short years risen to prominence for development, testing,
deployment and distribution of computer software. While there are
related approaches such as LXC or Singularity, Docker has become
synomymous with ``containerization''---the method of taking software
artefacts and bundling them in such a way that use becomes standardized
and portable across operationg systems. In doing so, Docker had
recognised and validated the importance of one very important thread
that had been emerging, namely virtualization. By allowing multiple
concurrent applications or services to run concurrently on one host
machine without any fear of interference between them, an important
scalability feature had been realized. But Docker improved on
virtualization by accessing the host system---generally Linux---through
a much thinner and smaller shim than a full operating system emulation
or virtualization. This makes for more efficient use of system resources
and allowed another order of magnitude in terms of scalability of
deployment. {[}Reference?{]} While Docker makes use of Linux kernel
features, its importance was large enough so that some required aspects
of running Docker have been added to other operating systems to support
Docker there more efficiently too. {[}Win10; reference?{]}

The key accomplishment of Docker is to make a ``bundled'' aggregation of
software available to any system running Docker, without requiring much
else from the host. This is a rather attractive, and novel, proposition
which has lead to widespread adoption and use of Docker in a variety of
domains. It provided also a natural match for ``cloud deployment'' which
runs (or at least appears to run) ``seamlessly'' without much explicit
reference to the underlying machine, architecture or operating system.
Using containers on cloud hosted servers is a natural fit also because
these containers can be deployed with very littler in terms of
dependencies on the host system.

For statistical computing and analysis centered around R, the Rocker
Project has provided a variety of Docker containers since its start in
2014 \citep{RJ-2017-065}. The Rocker Project provides several lines of
containeres spanning to from building blocks with R-release or R-devel
to containers with RStudio Server and Shiny Server. Also of note is a
series of ``versioned'' containers which match the R release they
contain with the \emph{then-current} set of packages via the MRAN
Snapshot views of CRAN. {[}Maybe at a later point add forward reference
to other topics building on Rocker described here.{]}

@@

\address{%
Dirk Eddelbuettel\\
University of Illinois at Urbana-Champaign\\
Department of Statistics\\ Illini Hall, 725 S Wright St\\ Champaign, IL 61820\\ ORCiD \href{https://orcid.org/0000-0001-6419-907X}{0000-0001-6419-907X}\\
}
\href{mailto:dirkd@eddelbuettel.com}{\nolinkurl{dirkd@eddelbuettel.com}}
